---
title: "Completion rates by game"
author: "Course example"
date: ""
output:
  pdf_document:
    latex_engine: xelatex
    fig_caption: true
  html_document: default
params:
  datafile: "steam_achievements.csv"
  minplayers: 100
---

```{r setup, include=FALSE}
library(ggplot2)
knitr::opts_chunk$set(
  echo = FALSE, cache = FALSE, error = FALSE,
  warning = TRUE, message = FALSE,
  fig.width = 6, fig.height = 3, out.width = "90%"
)
```

## Question and data

How do completion rates compare among games with at least the selected
number of players? A row represents one game; `completed` is the
number of its listed players who completed it. `achievements` counts
available achievements and is not the completion-rate denominator.

```{r prepare}
raw <- read.csv(params$datafile, stringsAsFactors = FALSE)
required <- c("game", "players", "completed", "achievements")
stopifnot(all(required %in% names(raw)), !anyNA(raw[required]))
counts <- raw[c("players", "completed", "achievements")]
stopifnot(all(vapply(counts, is.numeric, logical(1))))
stopifnot(all(is.finite(as.matrix(counts))))
stopifnot(all(counts >= 0), all(counts == floor(as.matrix(counts))))
stopifnot(all(raw$players > 0), all(raw$completed <= raw$players))
stopifnot(!anyDuplicated(raw$game), all(nzchar(raw$game)))
stopifnot(length(params$minplayers) == 1,
          is.finite(params$minplayers), params$minplayers >= 0)
cleaned <- raw[raw$players >= params$minplayers, ]
cleaned$rate <- cleaned$completed / cleaned$players
cleaned <- cleaned[order(-cleaned$rate, cleaned$game), ]
```

## Coverage and comparison

The threshold is at least `r params$minplayers` players per game.
This retains `r nrow(cleaned)` of `r nrow(raw)` games.

```{r table}
shown <- cleaned[c("game", "players", "completed", "achievements", "rate")]
knitr::kable(shown, row.names = FALSE, digits = 3)
```

The completion rate divides `completed` by `players`. Counts and rates
describe different aspects of these records.

```{r rates, fig.cap="Completion rates for eligible games."}
if (nrow(cleaned) == 0) {
  cat("No games meet this threshold. There are no rates to plot.\n")
} else {
  cleaned$game <- factor(cleaned$game, levels = rev(cleaned$game))
  print(ggplot(cleaned, aes(rate, game)) +
    geom_col(fill = "#D4A017") +
    scale_x_continuous(limits = c(0, 1)) +
    labs(x = "Completed / players", y = "Game") +
    theme_minimal(base_size = 11))
}
```

```{r players, fig.cap="Player counts for eligible games."}
if (nrow(cleaned) > 0) {
  print(ggplot(cleaned, aes(players, game)) +
    geom_col(fill = "#777777") +
    labs(x = "Number of players", y = "Game") +
    theme_minimal(base_size = 11))
}
```

Filtering changes coverage, not the underlying per-game rate. When games
qualify, the player-count plot makes their denominators visible.

## Reproduction record

```{r versions}
cat("Input", params$datafile, "\n")
cat("Input MD5", unname(tools::md5sum(params$datafile)), "\n")
cat("R", as.character(getRversion()), "\n")
for (pkg in c("ggplot2", "knitr", "rmarkdown")) {
  cat(pkg, as.character(packageVersion(pkg)), "\n")
}
```
