---
title: "Iris sepal lengths"
author: "Course example"
date: ""
output:
  pdf_document:
    latex_engine: xelatex
    fig_caption: true
  html_document: default
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
knitr::opts_chunk$set(
  echo = FALSE, cache = FALSE, error = FALSE,
  warning = TRUE, message = FALSE,
  fig.width = 6, fig.height = 3, out.width = "90%"
)
dat <- iris
```

## Question and data

How do sepal lengths vary within and between the three species in this
sample? Iris contains 150 flowers, 50 of each species, with measurements
in centimetres.

The mean sepal length is `r sprintf("%.2f", mean(dat$Sepal.Length))` cm,
with sample standard deviation `r sprintf("%.3f", sd(dat$Sepal.Length))` cm.

## A compact summary

```{r summary}
summary_table <- dat |>
  group_by(Species) |>
  summarise(
    n = n(), Sepal_cm = mean(Sepal.Length),
    Petal_cm = mean(Petal.Length), .groups = "drop"
  )
knitr::kable(summary_table, row.names = FALSE, digits = 3)
```

The table reports means in centimetres. Each species has the same sample
size, but a mean alone does not describe the within-species spread.

## Two views of the variation

```{r histogram, fig.cap="Sepal length distribution."}
edges <- seq(4.25, 8.25, length.out = 11)
ggplot(dat, aes(Sepal.Length)) +
  geom_histogram(
    breaks = edges, closed = "left",
    fill = "#D4A017", colour = "white"
  ) +
  labs(x = "Sepal length (cm)", y = "Number of flowers") +
  theme_minimal(base_size = 11)
```

Figure 1 uses 0.4 cm bins from 4.25 to 8.25 cm. The pooled histogram
combines species. Its bin edges avoid the recorded
0.1 cm measurement grid, so no observation lies on an internal edge.

```{r boxplot, fig.cap="Sepal lengths by species."}
ggplot(dat, aes(Species, Sepal.Length)) +
  geom_boxplot(fill = "#D4A017", width = 0.55) +
  labs(x = "Species", y = "Sepal length (cm)") +
  theme_minimal(base_size = 11)
```

Figure 2 uses linear quartiles and 1.5-IQR whiskers.
Setosa has the lowest sample median, while the species' ranges overlap.
The box plots reveal variation hidden by the table. They do not establish
that species alone causes a difference in sepal length.

## Reproduction record

Data are the built-in R `iris` dataset. The histogram uses the explicit
edges above; ggplot2 uses linearly interpolated quartiles for the boxes.
No random operation or cached analysis is used.

```{r versions}
cat("R", as.character(getRversion()), "\n")
for (pkg in c("ggplot2", "dplyr", "knitr", "rmarkdown")) {
  cat(pkg, as.character(packageVersion(pkg)), "\n")
}
```
